---
title: "HW2"
author: "Andrew Liang"
date: "9/11/2020"
output: pdf_document
---

**3.2)**
```{r}
y <- 57
n <- 100
theta0 <- seq(.1, .9, by = .1)
n0 <- c(1,2,8,16,32)

mat <- matrix(0, length(n0), length(theta0)) #create empty matrix to store prob values for post. beta distribution
rownames(mat) <- c(1,2,8,16,32)
colnames(mat) <- c(seq(.1, .9, by = .1))
#compute a and b given all combination theta0 and n0 values
for(i in 1:length(n0)){
  for (j in 1:length(theta0)){
    a = n0[i] * theta0[j]
    b = n0[i] * (1 - theta0[j])
#with our prior being a beta dist. for theta, our posterior is also beta(a+y,b+n-y) for theta 0 <= theta <= 1
mat[i,j] <- 1 - pbeta(.5, a + y, b + n - y)}} #posterior probs that theta >0.5 for each value of a and b

print(a)
print(b)
print(mat)

contour(n0, theta0, mat, xlab = 'n0', ylab = 'theta0', levels = seq(.1,1, by =.05))
```

From the output above, we can see that the values of a and b are 28.8 and 3.2 respectively for all the combinations of $\theta_0$ and $n_0$. The values for $Pr(\theta>0.5|\sum Y_i = 57)$ are also computed in the matrix above.

Based on the data that $\sum_{i=1}^{100} Y_i = 57$, the matrix probability, and posterior beta, we can see that the individuals who expect a prior $\theta<0.5$ would have around a 0.9 certainty that the posterior $\theta$ is greater than 0.5. As the sample size increases however, the posterior belief for these individuals decreases. For those who expect the prior $\theta>0.5$, they have around a 0.95 certainty that the posterior probability is greater than 0.5 regardless of the prior sample size.

**3.3a)**
```{r}
#95% Posterior CI's for part a)
CIa.theta <- qgamma( c(.025,.975), 237,20); CIa.theta
CIb.theta <- qgamma( c(.025,.975), 125,14); CIb.theta
```
**3.3b)**
```{r}
n0 <- c(seq(1,50, by = 1))

#posterior expectation of theta b given n0 values
expect.b <- (12*n0+113)/(n0+13); expect.b
plot(expect.b, xlab = "n0", ylab = "Posterior expectation of theta B")
```
We can see the posterior expected values of $\theta_{B}$ given $n_{0}$ values from the plot above. Since the expected value of $\theta_{A}$ is around 11.8, in order for the posterior expectation of $\theta_{B}$ to be close to that of $\theta_{A}$, we need to have the prior belief that $n_{0}$ is larger than 40 so that the posterior mean of $\theta_{B}$ can be close to the posterior mean of $\theta_{A}$; it seems that the expectation of $\theta_{B}$ reaches the expectation of $\theta_{A}$ asymptotically from the plot.

**c)**
Since it is stated that mice B and related to mice A, then it would not make sense to state that $\theta_{A}$ and $\theta_{B}$ are independent. Because the prior information about mice B comes from information from mice A, it wouldn't make sense to have $p(\theta_{A},\theta_{B}) = p(\theta_{A}) * p(\theta_{B})$. Instead, we should use $p(\theta_{A},\theta_{B}) = p(\theta_{A}) * p(\theta_{B}|\theta_{A})$.

**3.9a)**

A conjugate prior of the Galenshore distribution can also be the Galenshore distribution itself, although it seems that we can also use the Gamma distribution.
```{r}
dgalenshore <- function(x, a, theta){
  if (any(x <= 0) || any(a <= 0) || any(theta <= 0))
    stop("Invalid parameters")
  else {
    return((2/gamma(a)) * theta^(2*a) * x^(2*a-1) * exp(-theta^2 * x^2))
  }
}

x <- seq(0.01, 3, 0.01)
plot(x, dgalenshore(x, 1, 1), type = "l", xlab = "y", ylab = "Galenshore Density", lwd = 2, ylim = c(0,2))
lines(x, dgalenshore(x, 2,1), type = "l", col = "blue", lwd = 2, lty = 2)
lines(x, dgalenshore(x, 1,2), type = "l", col = "red", lwd = 2, lty = 3)
lines(x, dgalenshore(x, 2,2), type = "l", col = "purple", lwd = 2, lty = 4)
legend("topright", c("Galen(1,1)","Galen(2,1)","Galen(1,2)","Galen(2,2)"), col = c("black","blue","red","purple"), lwd = c(2,2,2,2), lty = c(1,2,3,4))
```
